Enabling GPU Virtualization in Cloud Environments

### 2015-10-23 paper.tex
state-of-art revisado

\section{\uppercase{Background}}
\label{sec:background}
\subsection{The rCUDA Framework}
\label{sec:rcuda}
{rCUDA}~\cite{toniparco} is a middleware that enables transparent access
to any NVIDIA GPU device present in a cluster from all compute
nodes. The GPUs can be accessed and shared between nodes, and a single node can use all these graphic accelerators
as if they were local.
These features are focused in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
resource, space, and energy~\cite{energy14} requirements.
rCUDA is structured following a client-server distributed
architecture: the client middleware is allocated in the same cluster node where the application demanding GPGPU
acceleration services is executed, providing a transparent replacement for the
native CUDA libraries. Furthermore, the server middleware is executed in the
cluster nodes from which the actual GPUs provide the requested GPGPU service.
To support a concurrent scenario where GPUs are shared between
processes\slash nodes, {rCUDA} manages separate device contexts for
each client application.

The {rCUDA} 5.0 client exposes the same interface as the regular NVIDIA
CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
APIs as well as other commonly used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
Therefore, applications are not aware of the fact that they are being executed
on top of a virtualization layer.
With the aim to be updated with new GPU programming models, {rCUDA} also supports
directive-based models such as OmpSS~\cite{repara15} and OpenACC~\cite{cluster15}.

The integration of remote GPGPU virtualization with global
resource schedulers such as SLURM~\cite{sbacpad14} completes this powerful
technology, making accelerator-enabled clusters more flexible and
energy efficient.

\subsection{Amazon Web Services}
\label{sec:aws}
AWS~\cite{aws} is a public cloud computing provider
and it is composed by several cloud services such as 
 cloud-based computation, storage and other functionality 
that enables organizations and/or individuals to deploy
services and applications on an on-demand basis. 
These services replace IT infrastructure for most companies and provide agility and instant elasticity matching 
perfectly with their software requirements.

From the point of view of HPC, AWS offers high performance facilities including 
instances equipped with GPUs and high performance network interconnection which 
are common resources used in this research field.

@@

OpenStack~\cite{OpenStack} is a cloud operating system that provides Infrastructure as a Service (IaaS). 
It is able to control large pools of compute, storage, and networking resources throughout a 
datacenter. All these resources are managed through a dashboard or an API that gives administrators 
control while empowering their users to provision resources through a web interface
or a command-line. 
It supports most available hypervisors and handles provisioning 
and life-cycle management of VMs.

The OpenStack architecture offers a flexibility to create a custom cloud, with no proprietary hardware
or software requirements and the ability to integrate with legacy systems and third party technologies.

From the HPC perspective, OpenStack offers high performance computing (HPC) configurations using
multiple supported hypervisors and different hardware architectures. However, despite to be able to
specify some extra parameters for special resources like GPUs, choosing a host based on the existence
of a GPU is currently unsupported in OpenStack~\cite{OpenStackGPU}.

\section{\uppercase{State of Art}}
\label{sec:state}
Our solutions is based on GPU virtualization and sharing resources in order
to reach a fair balance between supply and demand. 
Several attempts of achieving what we are pursing have been
made, but none of them are so ambitious as our project.

On
the one hand the work done in~\cite{younge2013enabling} allows the VM managed
by the hypervisor Xen to access the GPUs in the physical
node, with the implied limitations of restricting the nodes not
to use more GPUs than the hosted in the machine and not
to share idle GPUs to other machines. The solution presented 
by the project gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and make
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor used, so
does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. 
This one continues without being integrated in a Cloud Computing Manager, though its main
drawback is that the code of the applications must be modified
in order to be run in their virtual-GPU environment. A runtime component to provide abstraction
and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
to isolate and share GPUs in a cluster for a set of applications. Our proposal uses a different 
approach employing virtual machines and a cloud computing environment to allow to share GPUs in a flexible way.
On the other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
it is only focussed on compute intensive HPC applications.

The main idea of our proposal goes further and apart from
bringing solutions for all kind of HPC applications, it is aimed
to boost the cluster flexibility in the use of GPUs.

>>>>

\section{\uppercase{Background}}
\label{sec:background}
\subsection{The rCUDA Framework}
\label{sec:rcuda}
{rCUDA}~\cite{toniparco} is a middleware that enables transparent access
to any NVIDIA GPU device present in a cluster from all compute
nodes. The GPUs can be accessed and shared between nodes, and a single node can use all the graphic accelerators
as if they were local.
These features focus in attaining higher accelerator utilization rates in the overall system while simultaneously reducing
resource, space, and energy requirements~\cite{energy14}.
rCUDA is structured following a client-server distributed
architecture: the client middleware runs in the same cluster node where the application demanding GPGPU
acceleration services is executed, providing a transparent replacement for the
native CUDA libraries. Furthermore, the server middleware is executed in the
cluster nodes from which the actual GPUs provide the requested GPGPU service.
To support a concurrent scenario where GPUs are shared between
processes\slash nodes, {rCUDA} manages separate device contexts for
each client application.

The {rCUDA} 5.0 client exposes the same interface as the regular NVIDIA
CUDA 6.5 release~\cite{cuda65}, including the runtime and driver
APIs as well as other commonly used libraries such as cuBLAS, cuFFT, cuSparse or cuRand.
Therefore, applications are not aware of the fact that they are being executed
on top of a virtualization layer.
To deal with new GPU programming models, {rCUDA} has been recently extended to accommodate 
directive-based models such as OmpSS~\cite{repara15} and OpenACC~\cite{cluster15}.

The integration of remote GPGPU virtualization with global
resource schedulers such as SLURM~\cite{sbacpad14} completes this powerful
technology, making accelerator-enabled clusters more flexible and
energy-efficient.

\subsection{Amazon Web Services}
\label{sec:aws}
AWS~\cite{aws} is a public cloud computing provider, 
composed of several services, such as 
 cloud-based computation, storage and other functionality, 
that enables organizations and/or individuals to deploy
services and applications on an on-demand basis. 
These services replace local IT infrastructure for companies and provide agility and instant elasticity matching 
perfectly with their software requirements.

From the point of view of HPC, AWS offers high performance facilities via 
instances equipped with GPUs and high performance network interconnection, which 
are common resources leveraged in this research field.

@@

OpenStack~\cite{OpenStack} is a cloud operating system that provides Infrastructure as a Service (IaaS). 
It is able to control large pools of compute, storage, and networking resources throughout a 
datacenter. All these resources are managed through a dashboard or an API that gives administrators 
control while empowering their users to provision resources through a web interface
or a command-line interface.  
OpenStack supports most available hypervisors and handles provisioning 
and life-cycle management of VMs.

The OpenStack architecture offers flexibility to create a custom cloud, with no proprietary hardware
or software requirements and the ability to integrate with legacy systems and third party technologies.

From the HPC perspective, OpenStack offers high performance configurations using
multiple supported hypervisors and different hardware architectures. Despite in OpenStack is possible to 
specify some extra parameters requesting special resources, like GPUs, choosing a host based on the existence
of a GPU is currently unsupported~\cite{OpenStackGPU}.

\section{\uppercase{State-of-the-Art}}
\label{sec:state}
Our solutions to the defficiency exposed in the previous section is based on GPU virtualization, sharing resources in order
to attain a fair balance between supply and demand. 
While several efforst with the same goal have been made in the past, as exposed next, 
none of them is so ambitious as ours.

The work in~\cite{younge2013enabling} allows the VM managed
by the hypervisor Xen to access the GPUs in the physical
node, with the implied limitations that a node cannot 
use more GPUs than those hosted in the machine, and idle GPUs can not be shared with other machines. The solution presented 
by gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and makes
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor and, so
does its performance. Another similar solution is presented in gCloud~\cite{diab2013dynamic}. 
iWhile this solution is not yet integrated in a Cloud Computing Manager, its main
drawback is that the code of the applications must be modified
in order to be run in the virtual-GPU environment. A runtime component to provide abstraction
and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
to isolate and share GPUs in a cluster for a set of applications. 
The work introduced in~\cite{jungpgpu} is more mature, however,
it is only focussed on compute-intensive HPC applications.

Our proposal goes further and, in addition to 
bringing solutions for all kind of HPC applications, it is aimed
to boost flexibility in the use of GPUs.
